### 2015-04-04 be.tex
appendix B4

Here we derive an expression for the Bernoulli parameter, \be
\mathcal{B} \equiv \ke +\gammaf \frac{p}{\rho}+\Phi, \ee assuming a
stellar density profile $\rhostar\sim r^{-1-\Gamma}=r^{-\delta}$
characteristic of that inside the break radius $\rb$.

Basic manipulation of the equations of hydrodynamics
(eqs.~[\ref{eq:drhodt}]-[\ref{eq:dsdt}]) results in the following
relationship between the Bernoulli parameter and the net heating rate
$\propto q \vw^{2}$,
\begin{align}
\frac{1}{r^2} \frac{d}{dr} \left(r^2 \rho v \mathcal{B} \right) = q \kew + q \Phi = \frac{3}{1+\delta}\frac{G
  \Mbh}{r}+\sigma_0^2+v_w^2 + q \Phi,
\label{eq:enCons}
\end{align}
where the first and second terms on the right hand side correspond,
respectively, to the black hole and stellar contributions to the
velocity dispersion.

Integrating both sides of equation~\eqref{eq:enCons} from the
stagnation radius $r = \rs$ outwards,
\begin{align}
  r^2 \rho v \left(\ke+\Phi+\gammaf \frac{p}{\rho} \right)= \int_{\rs}^{r}
    r^2 \left(q \kew + q\Phi\right) dr
    \label{eq:enConsInt}
\end{align}
From mass continuity (eq.~[\ref{eq:drhodt}]) the mass flux can be
written

\begin{equation}
 f\equiv r^2 \rho v = q_o \rs^3 \left[\frac{x^{3-\delta}-1}{3-\delta}\right],
 \label{eq:massFlux}
 \end{equation}

where $x \equiv r/\rs$ and we have used the fact that the mass
 source term $q = q_o x^{-\delta}$ is proportional to the stellar
 density.  Likewise, the gravitational potential $\Phi$ is given by
\begin{align}
\Phi \simeq -\frac{G \Menc}{r} -4 \pi G \int_{r}^{\rb} \rhostar(r') r' dr',
\label{eq:Phi}
\end{align}
where by truncating the integral at $r = \rb$ we are neglecting the
contribution to the potential of the stars outside of the break
radius.

Combining equations~\eqref{eq:enConsInt} and \eqref{eq:massFlux} with
the result of equation \eqref{eq:Phi}, we obtain
\begin{align}
  &\mathcal{B} = -U_o\left[\frac{\rs/\rb}{2-\delta}-\frac{x^{5-2\delta}-1}{x^{3-\delta}-1}\frac{3-\delta}{(5-2\delta)(2-\delta)}\right]\nonumber\\
  &+\frac{v_w^2}{2}+\frac{\sigma_0^2}{2}- \frac{G \Mbh}{r_s}
  \frac{1-2\delta}{2(\delta+1)}
  \frac{3-\delta}{2-\delta}\frac{x^{2-\delta}-1}{x^{3-\delta}-1}
  \nonumber\\
  &-\frac{G M_\star}{r_s}
  \frac{x^{5-2\delta}-1}{x^{3-\delta}-1}\frac{3-\delta}{5-2\delta},
\end{align}
where $U_o=4 \pi G \rhostar|_{\rs} \rs^2$.  This can be rewritten as
\begin{multline}
  \mathcal{B} =\frac{G \Mbh}{\rs} \biggl[ \frac{3}{2} \zeta^2
  w^{\frac{1}{3 -\delta}}
  -\frac{(1-2\delta)(3-\delta)}{2(\delta+1)(2-\delta)}  \frac{x^{2  -\delta}-1}{x^{3-\delta}-1}\\
  -\frac{\rb}{\rsoi} w^{\frac{\delta-1}{3-\delta}} \frac{3 -\delta}{2
    -\delta} -w
  \frac{(2-\delta)(3-\delta)-(3-\delta)^{2}}{(5-2\delta)(2-\delta)}
  \frac{x^{5-2\delta}-1}{x^{3-\delta}-1} \biggr],
\label{eq:enthAnal}
\end{multline}
where $\zeta \equiv \sqrt{v_w^2+\sigma_0^2}/\sigma_0$ and $w\equiv
(\rs/\rsoi)^{3\delta}$, and $\delta$ is the power law slope of the
stellar density profile inside of the break radius, $\rb$.

Analysis of equation (\ref{eq:enthAnal}) results in the following.  When the rate of external heating is high ($\zeta \gg 1$), then $\mathcal{B} > 0$ at all radii, requiring an outflow.  As $\zeta$ is decreased, there exists a critical value of $\zeta = \zeta_{c}$ below which $\mathcal{B}$ becomes negative, first near the stagnation radius and then at larger radii as $\zeta$ is reduced further.  If $\zeta<\zeta_c$, then we have $\mathcal{B} > 0$ at the break radius, allowing an outflow with positive asymptotic kinetic energy, only if $\rs$ exceeds a minimum value, $r_{\rm s,min}$.

The value of $\zeta_c$ is an increasing function of $r_{b}/r_{\rm soi}$ and is larger for core galaxies than cusp galaxies, since a greater heating rate is required to unbind gas from a deeper stellar potential (Fig.~\ref{fig:zetaCrit}).  Specifically, we find $\zeta_c\sim (\rb/\rsoi)^{0.5(1-\Gamma)}$ for $\rb\gsim\rsoi$ (see the top panel of
Fig.~\ref{fig:zetaCrit} for the case of core galaxies). Likewise $(r_{\rm s,min}/r_{b})$ is an increasing function of $\zeta/\zeta_{c}$, approaching unity for $\zeta < 0.86\zeta_c$, as shown in the
bottom panel of Fig.~\ref{fig:zetaCrit}.  In practice, as $\zeta \rightarrow 0.86\zeta_c$ we find that the stagnation radius moves its value predicted by equations (\ref{eq:stag_analytic}), (\ref{eq:densSlope}) to the break radius.

\begin{figure}
\includegraphics[width=\columnwidth]{zetaCrit.pdf}
\caption{\label{fig:zetaCrit} {\emph Top panel:} Critical
  $\zeta_c=\sqrt{v_w^2+\sigma_0^2}/\sigma_0$ for core ($\Gamma=0.1$)
  galaxies. If $\zeta<\zeta_c$ outflows are only possible if the ratio
  of the stagnation radius to the influence radius exceeds a minimum
  value. Plotted as a function of the ratio of the break radius to the
  influence radius ($\rb/\rsoi$) {\emph Bottom Panel:} Mininmum value
  for ratio of the stagnation radius to the break radius, $\rs/\rb$
  for core galaxies ($\Gamma=0.1$) derived by requiring the Bernoulli
  parameter of the gas to remain positive out to the break radius
  $\rb$. This is calculated numerically from
  equation~\eqref{eq:enthAnal}.}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ms"
%%% End: